---
title: "Assignment_7"
author: "Weronika Lepiarz"
format: html
editor: visual
---

**Task 1:** Load the “iris” dataset and transform it to longer.

Uploading needed libraries.

```{r}
library(tidyr)
library(dplyr)
```

Loading "iris" dataset.

```{r}
data(iris)
head(iris)
```

Transforming data to longer.

```{r}
iris_long <- iris %>%
  pivot_longer(cols = -"Species", names_to = "Trait", values_to = "Size")
head(iris_long)
```

cols = - "Species" means EVERY column beside Species

**Task 2.1:** What is the **mean** **length** and **width** of the **sepal** and **petal** of each species?

```{r}
iris_mean <- iris_long %>%
  group_by(Species, Trait) %>%
  summarise(Mean_Size = mean(Size))
iris_mean
```

-   group_by(Species, Trait) - groups data by species and trait

-   summarise(Mean_Size = mean(Size)) - calculate mean value (from Size column) for each group

**Task 2.2:** What is the **largest** and **widest**,  **petal** and **sepal,**  of **virginica** and **versicolor** species?

```{r}
iris_max <- iris_long %>%
  filter(Species %in% c("virginica", "versicolor")) %>%
  group_by(Species, Trait) %>%
  summarise(Max_Size = max(Size))
iris_max
```

-   filter(Species %in% c("virginica", "versicolor")) - filters and keeps only Species as virginica and versicolor

-   group_by(Species, Trait) - groups data by species and trait

-   summarise(Max_Size = max(Size)) - gets maximum value (from Size column) for each group

**Task 2.3:** For each species, for sepal and petal width and length, how many observations are above the mean?

```{r}
iris_above_mean <- iris_long %>%
  left_join(iris_mean, by = c("Species", "Trait")) %>%
  filter(Size > Mean_Size) %>%
  group_by(Species, Trait) %>%
  summarise(Count_Above_Mean = n())
iris_above_mean
```

-   left_join(iris_mean, by = c("Species", "Trait")) - adds mean row to every row

-   filter(Size \> Mean_Size) - filters and keeps rows where Size is greater than Mean_Size

-   group_by(Species, Trait) - groups data by species and trait

-   summarise(Count_Above_Mean = n()) - counts how many rows passed the filter
